---
title: "DATA2002 Assignment"
author: "Jing Yang \ 480110301"
output:
  bookdown::html_document2: # note that this requires bookdown
    toc: yes
    toc_depth: 3
    toc_float: yes
    code_folding: hide
    number_sections: true
    theme: paper
    highlight: kate
    css: style.css
---

# Executive Summary

# Introduction

```{r setup, include=TRUE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(skimr)
library(visdat)
library(gendercodeR)
library(kableExtra)
library(coin)

raw = read_csv("DATA2X02 class survey 2020 (Responses) - Form responses 1.csv")

x = raw %>% janitor::clean_names()
```

# Data Cleaning

```{r}
colnames(x) = stringr::str_replace(string = colnames(x),
                                   pattern = "what_is_your_",
                                   replacement = "")
colnames(x) = stringr::str_replace(string = colnames(x),
                                   pattern = "on_average_how_many_hours_per_week_did_you_",
                                   replacement = "")

colnames(x)[2] = "covid_test"
colnames(x)[4] = "postcode"
colnames(x)[5] = "dentist"
colnames(x)[6] = "university_work"
colnames(x)[7] = "social_media"
colnames(x)[8] = "dog_or_cat"
colnames(x)[9] = "live_with_parents"
colnames(x)[10] = "exercising"
colnames(x)[12] = "asthma"
colnames(x)[13] = "paid_work"
colnames(x)[14] = "fav_season"
colnames(x)[16] = "height"
colnames(x)[17] = "floss_frequency"
colnames(x)[18] = "glasses"
colnames(x)[20] = "steak_preference"
colnames(x)[21] = "stress_level"

x = x %>% mutate(
  gender = gendercodeR::recode_gender(gender)
)
```

```{r}
detail = data.frame(Position=1:21, "Survey question"=colnames(raw),"Variable name"=colnames(x))
colnames(detail)=c("Position","Survey question","Variable name")
knitr::kable(detail,align="cll",caption = "Table of survey questions and corresponding variable names for data processing.") %>% kable_paper(full_width = F,html_font = "Cambria")
```

In Figure \@ref(fig:missingness) below we can see that some missingness exists and we should probably go back and remove the rows that are completely empty (apart from a timestamp). To do this you might consider using the `filter()` function or the more specific `drop_na()` function.

```{r missingness, fig.cap="Visualising the missingness in the data."}
visdat::vis_miss(x)+theme(axis.text.x = element_text(angle = 90))
x = x %>% drop_na()
```

# Questions

## Is this a random sample of DATA2002 students?

A random sample is a sample chosen from a larger set. Each individual is chosen randomly and entirely by enchange, such that each individual has the same proability of being chosen. This is a voluntary survey. No students are forced to answer the survey therefore every DATA2002 student has the same probability to complete the survey at any stage during the sampling process. Thus this is a random sample of DATA2002 students.

## What are the potential biases? Which variables are most likely to be subjected to this bias?

Voluntary Response Bias can occur in this research as sample members are self-selected volunteer. 

## Are there any questions that needed improvement to generate useful data?

The questions "How tall are you?" and "What is your shoe size?" shold be improved. Height and shoe size should have a natual limited range. However, In the responses of these questions, every extreme values exist and the standard deviation is quite large. This might because students answered questions with numeric responses in different units. For example, for the height respones less than 2.3, the unit used is obviously "meter" whereas the unit for height respones greater than 100 is centimeter.  When speccifying shoe sizes, students may refer to different sizing system (UK, US, Europe, Modopoint, etc.). 

To improve the questions, units should be specified in the questions so that numeric responses fall in reasonable ranges.The question "How tall are you?" can be changed to "What's your height in centimeter?" while the question "What is your shoe size?" can be changed to "What is your shoe size in UK sizing system?". 

## Does the number of COVID tests follow a Poisson distribution?

```{r results='hide'}
x %>% janitor::tabyl(covid_test) %>% 
  arrange(covid_test) %>% # arrange in descending order
  gt::gt() %>%
  gt::fmt_percent(columns = c("percent"),
                  decimals = 0) %>%
  gt::tab_source_note("Table: Summary of covid test proportions.")
```

As shown in Figure \@ref(fig:covidplot), the peak of the number of times students have been tested for COVID lies to the left with a long tail extending to the right. Therefore the data is significantly right-skewed. A goodness of fit test is performed to test whether these observations follow a Poisson distribution.

```{r covidplot, fig.cap="Distribution of how many times students have been tested for COVID."}
x %>% ggplot() + aes(x = as.factor(covid_test)) + 
  geom_bar() + scale_x_discrete(name = "Number of COVID test",limits = c("0","1","2","3","4","5","6","7","8","9","10"))

# p1 = x %>% ggplot(aes(x = covid_test)) + geom_bar()
```

There are 8 groups of the number of COVID test: 1,2,3,4,5,6 and 10. The corresponding observed frequencies are 101, 22, 8,3,1,2,1 and 1. $\lambda$ of the Poisson distribution is estimated to be 0.55.

```{r}
tab = as.data.frame(table(x$covid_test))
covid_count = tab$Freq
groups = as.numeric(as.character(tab$Var1))
n=sum(covid_count) # sample size
k = length(covid_count) # number of groups
(lam = sum(covid_count * groups)/n) # estimate the lambda parameter
```

**Hypothesis:** $H_0:$ the data come from a Poisson distribution vs $H_1:$ the data do not come from a Poisson distribution.

**Assumptions:** The expected frequencies, $e_i=np_i \geq 5$. Observations are independent.

The expected frequencies are 79.9, 44.2, 12.3, 2.3, 0.3, 0.03, 0.003 and 0. The expected frequencies $e_4, e_5, e_6, e_7, e_8$ are less than 5, therefore we conbine these 4 classes to satisfy assumption. 

```{r}
p = dpois(groups, lambda = lam) # obtain the p_i from the Poisson omf
(ey = n*round(p,5)) # calculate the expected frequencies
```

```{r}
ey >= 5 # check assumption e_i >= 5 not all satisfied
```

```{r}
(eyr = c(ey[1:3],sum(ey[4:8])))
```

However, after combining the 8 classes into 4 classes (i.e. $0, 1, 2, \geq 3$), the expected frequency of group $\geq 3$ is still less than 5. Therefore we combine the group "2" and "$\geq 3$". The final adjacent groups are "0","1" and "$\geq 2$". The corresponding expected frequencies are 79.9, 44.2 and 14.9 while the observed frequencies are 101, 22,16 respectively. 

```{r}
(eyr = c(ey[1:2],sum(ey[3:8])))
```

```{r}
(yr = c(covid_count[1:2],sum(covid_count[3:8])))
(pr = c(p[1:2],sum(p[3:8])))
```

```{r covidF, fig.cap="Distribution of the observed and expected frequencies for different number of COVID tests"}
xr = c("0","1",">=3") # group labels
par(mfrow=c(1,2), cex = 1.5)
data = data.frame("xr" = xr, "yr" = yr,"eyr" =eyr)

times = c("0","1",">=3")
data = data %>% 
  mutate(
    xr = factor(xr, levels = times)
  )

ob = ggplot(data, aes(x=xr,y=yr))+geom_bar(stat="identity")+labs(x="Number of COVID tests", y="Observed frequency")
ex = ggplot(data, aes(x=xr,y=eyr))+geom_bar(stat="identity")+labs(x="Number of COVID tests", y="Expected frequency")
gridExtra::grid.arrange(ob, ex, ncol = 2)
```

The expected frequencies are all greater than 5 so that the assumption $e_i \geq 5$ is met. As this is a random sampling research, the assumption of independence is met.

**Test statistic:**$T=\sum^k_{i=1}\frac{(Y_i-np_i)^2}{np_i}$.

Under $H_0$, T~$x_2^2$ approximately.

**Observed test statistic:** $t_0=16.86$
```{r}
kr = length(yr) #number of combined classes
(t0 = sum((yr - eyr)^2/eyr))
```

**P-value:** $P(T \geq t_0)=P(x_2^2\geq 16.86)=4e^{-5}$
```{r}
(pval = 1 - pchisq(t0, df = kr-1-1))
```

**Conclusion:** There p-value is $4e^{-5} < 0.05$. The $H_0$ is rejected therefore the number of COVID test follows a Poisson distribution.

## Perform other hypothesis tests. 

### Are men more likely to spend time on exercising compared to women?

From Figure \@ref(fig:exercising) , the means are different
the right plot removes outliers so that the patterns within boxes are more easily observed and compared. 
```{r exercising, fig.cap="Number of hours spent on exercising by different genders. The left boxplot shows all data including outliers while the right plot removes outliers. The means are shown by the blue dots in the right plots."}

x_gender = x %>% filter(gender!="non-binary")

p1 = x_gender %>% ggplot() + 
  aes(y = as.factor(gender), x = exercising) + 
  geom_boxplot() + 
  labs(y = "Gender", x = "Number of hours spent on exercising" ) +
  theme(legend.position = "none") + coord_flip()

p2 = x_gender %>% ggplot() + 
  aes(y = as.factor(gender), x = exercising) + 
  geom_boxplot(outlier.shape=NA) + 
  labs(y = "Gender", x = "Number of hours spent on exercising" ) +
  theme(legend.position = "none") + xlim(0, 15)+stat_summary(fun.y=mean, geom="point", shape=20, size=2, color="blue", fill="red") + coord_flip()

gridExtra::grid.arrange(p1, p2, ncol = 2)
```

```{r mean}
dat = x_gender%>% select(gender,exercising)
dat$gender = as.factor(dat$gender)

sum = dat %>% group_by(gender) %>%
      summarise(Mean = mean(exercising), 
      Median = median(exercising),
      SD = sd(exercising),
      Variance = var(exercising),
      n = n())

colnames(sum)=c("Gender","Mean","Median","SD","Variance","Number of students")
knitr::kable(sum,digits = 1, align="cccccc",caption = "Statistics of number of hours spent on exercising by different genders.") %>% kable_paper(full_width = F, position = "float_right",html_font = "Cambria")

x1 = x %>% mutate(social_media = case_when(
    social_media == "Youtube" ~ "Youtube",
    social_media == "youtube" ~ "Youtube",
    social_media == "YouTube" ~ "Youtube",
    social_media == "Wechat" ~ "Wechat",
    social_media == "WeChat" ~ "Wechat",
    social_media == "wechat" ~ "Wechat",
    social_media == "twitter" ~ "Twitter",
    social_media == "Twitter" ~ "Twitter",
    social_media == "Tiktok" ~ "Tiktok",
    social_media == "TikTok" ~ "Tiktok",
    social_media == "tiktok" ~ "Tiktok",
    social_media == "Tik Tok" ~ "Tiktok",
    social_media == "Snapchat" ~ "Snapchat",
    social_media == "reddit" ~ "Reddit",
    social_media == "Reddit" ~ "Reddit",
    social_media == "Pinterest" ~ "Pinterest",
    social_media == "no" ~ "None",
    social_media == "None" ~ "None",
    social_media == "Messenger" ~ "Messenger",
    social_media == "messenger" ~ "Messenger", 
    social_media == "linkedin" ~ "Linkedin",
    social_media == "Instagram" ~ "Instagram",
    social_media == "Instragram" ~ "Instagram",
    social_media == "Instangram" ~ "Instagram",
    social_media == "INSTAGRAM" ~ "Instagram",
    social_media == "instagram" ~ "Instagram",
    social_media == "insta" ~ "Instagram",
    social_media == "ins" ~ "Instagram",
    social_media == "Google+" ~ "Google+",
    social_media == "Facebook Messenger" ~ "Messenger",
    social_media == "Facebook messanger" ~ "Messenger",
    social_media == "facebook" ~ "Facebook",
    social_media == "Facebook" ~ "Facebook",
    social_media == "canvas discussion board" ~ "Canvas discussion board",
    social_media == "Club Penguin" ~ "Club Penguin",
    social_media == "Discord" ~ "Discord",
    social_media =="Ed" ~ "Ed",
    social_media =="bilibili" ~ "Bilibili"
  ) )
```

Table \@ref(tab:mean) shows that both mean and median number of hours spent on exercising are higher for males compared to females. The variance of males' exercising hour is also much greater than that of females.

Two hypothesis testing applying two-sample t-test and Mood's median test are performed to validate whether means and medians of exercising hour for different genders differ.

#### Two-sample t-test

The variances of two samples are different, violating the assumption of regular two-sample t-test of equal population variances. Thus, in stead of a regular two-sample t-test, we apply the Welch two-sample t-test which does not assume equal population variances.

**Hypothesis:** $H_0:$ Gender and number of hours spent on exercising are independent $(\mu_F=\mu_M)$ **vs** $H_1:$ Gender and number of hours spent on exercising are related (Not all equalities hold). 

**Assumptions:** $X_1,...,X_{n_x}$ are iid $N(\mu_X,\sigma^2),$ $Y_1,...,Y_{n_y}$ are iid $N(\mu_Y,\sigma^2)$ and $X_i$'s are independent of $Y_i$'s.

This is a random sample of DATA2002 students therefore obserations are supposed to be independent and identically distributed. As such, all assumptions are fulfilled.

Under $H_0,T$~$T_{n_x+n_y-2}$,
**Test statistics:** $T=\frac{\overline{X}-\overline{Y}}{\sqrt{\frac{S_x^2}{n_X}+\frac{S_y^2}{n_Y}}}$.

**Observed test statistic:** $t_0=\frac{92-46}{ \sqrt{\frac{7.6^2}{92}+\frac{2.9^2}{46}}}=2.58.$
```{r}
ny = sum[[1,5]] # female
nx = sum[[2,5]] # male

sy = sum[[1,4]]
sx = sum[[2,4]]

y_mean = sum[[1,2]]
x_mean = sum[[2,2]]
(t0 = (x_mean-y_mean)/sqrt(sx^2/nx+sy^2/ny))
```

**P-value:** $P(t_{n_x+n_y-2}\geq |2.58|)=0.01086.$
```{r}
female = x %>% filter(gender == "female") %>% select(exercising)
male = x %>% filter(gender == "male") %>% select(exercising)
t.test(male,female, alternative = "two.sided")
```

**Decision:** As p-value is small than 0.05, we reject $H_0$ and conclude that the mean number of hours spent on exercising by males is differrent from the mean exercising hours spent by females. Therefore gender and number of hours spent in exercising is related.

#### Median test

**Hypothesis:** $H_0:$ Gender and number of hours spent on exercising are independent $(\widetilde{x}= \widetilde{y})$ **vs** $H_1:$ Gender and number of hours spent on exercising are related $(\widetilde{x}\neq \widetilde{y})$.

**Assumptions:** $X_1,...,X_{n_x}$ are iid $N(\mu_X,\sigma^2),$ $Y_1,...,Y_{n_y}$ are iid $N(\mu_Y,\sigma^2)$ and $X_i$'s are independent of $Y_i$'s.

This is a random sample of DATA2002 students therefore obserations are supposed to be independent and identically distributed. As such, all assumptions are fulfilled.

Under $H_0,T$~$T_{(r-1)(c-1)}$ where r is the number of rows and c is the number of columns in contingency table,

**Test statistics:**

$T=\sum{\frac{(O-E)}{E}}=-3.0066$, where and O is the observed frequencie E is the expected frequencie in the contingency table.

**P-value:** $P(t_{(r-1)(c-1)}\geq |-3|)=0.002642$
```{r}
median_test(exercising ~ gender, data=dat)
```

**Decision:** As p-value is less than 0.05, we reject $H_0$ and conclude that the median exercising hours spent by males is different from that for females. Therefore exercising hour is related to gender. 

As discussed above by the two hypotheses, both mean and median exercising hours by different genders do not equal. Although both two-sample t-test and two-sample Mood's median test give the same result that exercising hour is related to gender, we regard the median test as the more reliable test compared to t-test in this case. Due to the existence of outliers, the two-sample t-test testing means may be affected by the extreme valeus while the Mood's median test is more robust to outliers and the shapes of the population distribution.

### Are gender and favorite social media related?

From Figure \@ref(fig:media), Instagram, Facebood and Reddit are the most popular social media within DATA2002 students. It appears that females and males favor different kinds of social media as the distribution of green and red bars are different. In particular, females are more likely to use Instagram while more males prefer other social media.

```{r media, fig.cap="DATA2002 students' favorite social media"}
media_f = as.data.frame(table(x1$social_media))
sort = arrange(media_f, Freq)
x1 %>% ggplot() + 
  aes(y = social_media,fill = gender, stat = "count")   +
  geom_bar() + 
  scale_y_discrete(name=" ", limits=as.character(sort$Var1)) +
  labs(y = "", x = "Count")+theme(plot.title = element_text(hjust = 0.5))
```

To perform a hypothesis testing, a contingency table with size of  $3 \times 18$ is obtained. From Table \@ref(tab:contingency), many expected frequencies in the contingency table are less than 5, obeying the assumption of Chi-square test. In addition, it is difficult to compute such a large contingency table by Fisher's exact test. Thus, we use an alternative method **Yates' continuity correction** to test in these small samples.

```{r contingency}
y.mat1 = table(x1$social_media,x1$gender)

r1 = nrow(y.mat1)
c1 = ncol(y.mat1)
yr1 = apply(y.mat1,1,sum) #rowSums(y.mat1)
yc1 = apply(y.mat1,2,sum) #colSums(y.mat1)
yr.mat1 = matrix(yr1, r1, c1, byrow = FALSE)
yc.mat1 = matrix(yc1, r1, c1, byrow = TRUE)
ey.mat1 = yr.mat1 * yc.mat1 / sum(y.mat1)
rownames(ey.mat1)=rownames(y.mat1)
colnames(ey.mat1) = colnames(y.mat1)
knitr::kable(ey.mat1, align = "lccc",caption ="Contingency table",digits = 1) %>%
  kable_classic(full_width = F, html_font = "Cambria",position = "float_right")
```

**Hypothesis:** $H_0:$ gender and favorite social media are independent ($p_{ij}=p_ip_j$ where $i=1,2,3$ and $j=1,2,3,...,18$) **vs** $H_1:$ gender and favorite social media are related (not all equalities hold).

**Assumptions:** Observations are independent.

As data is obtained form random sampling, we suppose the observations are independent. Thus the only assumption is met.

Under $H_0$, $T$ ~ $\sum^2_{i=1}\sum^2_{j=1}{\frac{(|Y_{ij}-e_{ij}|-0.5)^2}{e_{ij}}}$

**Test statistics:** $t_0=307.3$.

```{r}
res.yates1 = (abs(y.mat1 - ey.mat1)-0.5)^2/ey.mat1
(t0 = sum(res.yates1))
```

**P-value:** $P(T \geq t_0)=P(x_2^2\geq 307.3)=8.3e^{-69} \approx0$

```{r}
pchisq(t0,1,lower.tail=FALSE)
```

**Decision:** As p-value < 0.05, we reject $H_0$. Therefore gender and favorite social media are related.

# References

```{r, include = FALSE}
citation("tidyverse")
citation("janitor")
citation("visdat")
citation("gendercodeR")
citation("gt")
citation("skimr")
citation()
```

- Beaudry J, Emily Kothe, Felix Singleton Thorn and Rhydwyn McGuire (2020). gendercodeR: Recodes Sex/Gender Descriptions Into A Standard Set. R package version 0.0.0.9000. https://github.com/ropenscilabs/gendercoder
- Firke S (2020). janitor: Simple Tools for Examining and Cleaning Dirty Data. R package version 2.0.1. https://CRAN.R-project.org/package=janitor
- Iannone R, Joe Cheng and Barret Schloerke (2020). gt: Easily Create Presentation-Ready Display Tables. R package version 0.2.2. https://CRAN.R-project.org/package=gt
- R Core Team (2020). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. https://www.R-project.org/.
- Tierney N (2017). "visdat: Visualising Whole Data Frames." _JOSS_, *2*(16), 355. doi: [10.21105/joss.00355](https://doi.org/10.21105/joss.00355)
- Waring E, Michael Quinn, Amelia McNamara, Eduardo Arino de la Rubia, Hao Zhu and Shannon Ellis (2020). skimr: Compact and Flexible Summaries of Data. R package version 2.1.2. https://CRAN.R-project.org/package=skimr
- Wickham H et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, doi: [10.21105/joss.01686](https://doi.org/10.21105/joss.01686)
- Hao Zhu, Thomas Travison, Timothy Tsai, Will Beasley, Yihui Xie and  GuangChuang Yu (2020). kableExtra: Construct Complex Table with 'kable' and Pipe Syntax. R package version 1.2.1. https://cran.r-project.org
- Torsten Hothorn, Henric Winell, Kurt Hornik, Mark A. van de Wiel and Achim Zeileis (2019). coin: Conditional Inference Procedures in a Permutation Test Framework. R package version 1.3.1. https://cran.r-project.org