---
title: "DATA2002 Tut 4"
author: "Jing Yang \ 480110301"
date: "9/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE,warning=FALSE}
library(tidyverse)
```

# Exercise

## Personality type

**Hypothesis:** $H_0: p_{1j}=p_{2j}$ for $j=1,2,3,4,5$ **vs** $H_1:$ Not all equalities hold.

**Assumptions:** $e_{ij}=\frac{y_iy_j}{n}\geq 5$.
**Test statistic:** $t_0=\sum^2_{i=1}\sum^3_{j=1}{\frac{y_{ij}-e_{ij})^2}{e_{ij}}}=3.006$

**P-value:** $P(T\geq t_0)=P(x^2_2 \geq 3.006)=0.5568$

**Decision:** The p-value is greater than 0.05, there is not enough evidence to reject $H_0$. The data is consistent with $H_0$ therefore we conclude that personality types are homogeneous across Business and Social Science majors.

```{r}
counts = c(41, 52, 46, 61, 58, 72, 75, 63, 80, 65)
c_mat = matrix(counts, nrow = 2, byrow = TRUE)
colnames(c_mat) = c("Open", "Conscientious", "Extrovert", "Agreeable", 
    "Neurotic")
rownames(c_mat) = c("Business", "Social Science")
# c_mat
chisq.test(c_mat,correct = FALSE)
```

## Shcoking

### Fisher’s exact test
```{r}
counts = c(12,5,4,9)
y.mat = matrix(counts,nrow = 2, byrow = TRUE)
colnames(y.mat) = c("Wait together", "Wait alone")
rownames(y.mat) = c("High anxiety","Low anxiety")

fisher.test(y.mat)
```

### A chi-squared test for independence without a continuity correction

```{r}
chisq.test(y.mat, correct=FALSE)
```
### A chi-squared test for independence with a continuity correction

```{r}
chisq.test(y.mat, correct=TRUE)
```

### A chi-squared test for independence using a Monte Carlo p-value (i.e. using simulation).

```{r}
chisq.test(y.mat, simulate.p.value = TRUE, B=10000)
```

## Asbestos
```{r}
asbestos = matrix(c(310, 212, 21, 25, 7, 36, 158, 35, 102, 35, 0, 9, 17, 49, 51, 0, 0, 4, 18, 28), nrow = 5)
colnames(asbestos) = c("None", "Grade 1", "Grade 2", "Grade 3")
rownames(asbestos) = c("0-9", "10-19", "20-29", "30-39", "40+")
y = asbestos %>% as.data.frame() %>% 
  tibble::rownames_to_column(var = "years") %>% 
  tidyr::gather(key = grade, value = count, -years)
y$grade = factor(y$grade, levels = c("None", "Grade 1", "Grade 2", "Grade 3"), ordered = TRUE)
ggplot(y, aes(x = years, y = count, fill = grade)) + 
  geom_bar(stat = "identity") + 
  theme_bw(base_size = 16) + 
  scale_fill_brewer(palette = "Set1") + 
  labs(fill = "", y = "Count", x = "Occupational exposure (yrs)")
```
### Adapt the ggplot2 code above such that the y-axis is a proportion within each exposure length group. Does it look like there’s a relationship between the two variables?
```{r}
ggplot(y, aes(x = years, y = count, fill = grade)) + 
  geom_bar(stat = "identity",position="fill") + 
  theme_bw(base_size = 16) + 
  scale_fill_brewer(palette = "Set1") + 
  labs(fill = "", y = "Count", x = "Occupational exposure (yrs)")
```

It does loke like there's a reltationship between the two variables.

### Use the function chisq.test() to perform a standard chi-squared test of independence to determine whether there exists a statistically significant association between years of exposure to asbestos fibres and the severity of asbestosis that they were diagnosed with.

```{r}
chisq.test(asbestos)
```

### Use x = r2dtable(____) to randomly generate a contingency table with the same row and column totals as asbestos. Perform a chi-squared test and extract the test statistic using chisq.test(x[[1]])$statistic.

```{r}
r = rowSums(asbestos)
c = colSums(asbestos)
x = r2dtable(100,r,c)

chisq.test(x[[1]])
chisq.test(x[[1]])$statistic
```

### By using the r2dtable() function, perform a Monte-Carlo simulation to determine the p-value for the chi-squared test of independence. Generate 10,000 bootstrap resamples. Note: if doing this in an Rmd script, you might want to wrap your chisq.test(___)$statistic in suppressWarnings() so they don’t slow down your computer, e.g. suppressWarnings(chisq.test(___)$statistic). Plot a histogram of your Monte Carlo test statistics.

```{r}
chisq.test(x[[1]], simulate.p.value = TRUE, B=10000)
```

### Use the chisq.test() function to perform a Monte-Carlo simulation that obtains a p-value. Do so using 10,000 bootstrap resamples.

```{r}
chisq.test(asbestos, simulate.p.value = TRUE, B =10000)
```

# For after the lab

## IQ and Income

```{r}
mat = matrix(c(14,18,25,8,23,15),nrow = 3)
mat
```
```{r}
chisq.test(mat)
```

## Eating habits and living arrangements

```{r}
mat = matrix(c(2,25,23,146,7,124),nrow = 3)
chisq.test(mat)
```

### TV violence
```{r}
mat = matrix(c(8,12,21,18,15,7),nrow = 2)
chisq.test(mat)
```

