---
title: "Lab 10"
author: "Jing Yang \ 480110301"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

# Questions

```{r}
pollut = read_csv("https://raw.githubusercontent.com/DATA2002/data/master/pollut.txt")
glimpse(pollut)
```

1. Generate a pairs plot of the data using pairs() or the ggpairs() function from the GGally package.

```{r}
library(GGally)
ggpairs(pollut)
```
2. Perform a multiple regression of ozone on the other variables using lm().

```{r}
full = lm(O~.,data=pollut)
round(summary(full)$coef,3)
```

3. Does it look like any variables can be dropped from the model? If you were doing backwards selection using the drop1() function which would you drop first? Write down a the workflow for a formal hypothesis test to see if the coefficient for insolation is significantly different to zero.

`In` is dropped.
```{r}
drop1(step.back.aic,test="F")
```



4. Rather than dropping variables using their individual p-values, we can instead consider using an information criterion. Use the step() function to perform selection using the AIC starting from the full model.
```{r}
library(sjPlot)
step.back.aic = step(full,
                     direction = "backward",
                     trace = FALSE)
(lm_b = drop1(step.back.aic,test="F"))
```


5. Write down the fitted model for the model selected by the step-wise procedure.

```{r}
# remotes::install_github("datalorax/equatiomat")
# library(equatiomatic)
# extract_eq(lm2)
library(equatiomatic)
extract_eq(lm_b)
```

6. Check the linear regression assumptions for the stepwise model.

```{r}
full$fitted.values
pollut = pollut %>%
mutate(
lfitted = lm_b$fitted.values,
lresid = lm_b$residuals
)

library(ggfortify)
p2 = ggplot(pollut, aes(x = Temp,
y = lresid)) +
geom_point(size = 3) +
theme_classic(base_size = 30) +
labs(x = "Temperature (°F)",
y = "Residual") +
geom_hline(yintercept = 0)
p2
```

