---
title: "DATA Lab 5"
author: "Jing Yang \ 480110301"
date: "9/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
bac = c(12.3, 12.7, 12.6, 13.1, 13.2, 12.8, 13.1, 12.9, 13.1, 12.4, 
    13.6, 12.7, 12.6, 13.1, 12.4, 12.6, 13.3, 12.6, 12.4, 13.1, 12.9, 
    12.6, 12.7, 12.5, 12.4, 12.4, 12.6, 12.7, 12.4, 12.9)
n = length(bac)
xbar = mean(bac)
s = sd(bac)
c(n, xbar, s)
```

# Blood alcohol readings

**a.**

**Hypothesis:** $H_0: \mu=12.6$ vs $H_1: \mu \neq 12.6$


**b.**
Assumptions: $X_i$ are *iid* varialbes and follow $N(\mu,\sigma^2)$.

The assumption is violated as the boxplot and Normal Q-Q plot shows right-skewness of the data.
```{r}
boxplot(bac,horzontal = TRUE, xlab="Blood alcohol")

qqnorm(bac);qqline(bac)
```

**c.** 29 degree of freedom


**d.** 

Test statistic: $T=\frac{\overline{X}-\mu_{0}}{S/\sqrt{n}}$. Under $H_0$, T~$t_{n-1}$.

Observed test statistic: $t_0=\frac{12.76-12.6}{0.32/\sqrt{30}}=2.64$

**e** 

Critical value: $t_{0.05}=2.045$. i.e. reject if $t_0$ is greater than $2.045$

```{r}
qt(c(0.9, 0.95, 0.975), 30) # calculating T0
```

Decision: the observed test statistic $t_0=2.64$ is greater than 2.045, so reject $H_0$. Thus we conclude that the Analyzer GTE-10 is faulty.

**f**
```{r}
t.test(bac, mu = 12.6, alternative = "two.sided", conf.level = 0.95)
```

# Life satisfaction

```{r message=FALSE} 
library("tidyverse")
Young = c(24, 26, 40, 29, 29, 41, 32, 19, 23, 25, 37, 31, 31, 29, 24, 
          42, 32, 13, 33, 25, 20, 26, 20, 23, 23, 15, 34, 29, 20, 38)
Old = c(27, 26, 45, 34, 34, 45, 36, 20, 22, 24, 35, 31, 26, 41, 31, 
        37, 31, 12, 38, 26, 22, 27, 21, 31, 23, 24, 27, 33, 22, 40)

dat = tibble(
  satisfaction = c(Young, Old),
  age = c(rep("Young", length(Young)),
          rep("Old", length(Old)))
)

dat %>% 
  group_by(age) %>%
  summarise(
    n = n(),
    mean = mean(satisfaction),
    sd = sd(satisfaction)
  )
```

```{r}
par(mfrow = c(1,2))
boxplot(Young,Old,names = c('Young','Old'))
boxplot(Old - Young)
axis(1, at = 1, labels = 'Old - Young')
```

**a.** A hypothesis with two-sample Mann-Whitney U test should be performed.

**b.** $H_0: \mu_x=\mu_y$ vs $H_1:\mu_x \neq \mu_y$

**c.** Assumptions: $x_i$ and $y_i$ are independent and follow the ame kind of distribution but differ by a shift.

```{r warning=FALSE}
wilcox.test(Old,Young,alternative = 'two.sided', correct = FALSE)
```

**d.** Test statistic: $W=R_1+R_2+...R_{n_x}$ (the sum of the ranks of observations in old adults. Under $H_0$, W~$WRS'(30,30)$.

Observed test statistic: $w=r_1+r_2+...+r_{n_x}=513.5$

**e.** Decision: as p-value is 0.34, which is greater than 0.05, we do not reject $H_0$ and conclude that there is no difference of life satisfaction between old adults and young adults.

**f.** 
```{r}
t.test(Old, Young, alternative = 'two.sided')
```

```{r}
paired_dat = tibble(Young, Old) %>%
  mutate(
    Difference = Old - Young
  )
glimpse(paired_dat)

statistics = paired_dat %>% 
  summarise(
    n = n(),
    mean_old = mean(Old),
    mean_young = mean(Young),
    mean_diff = mean(Difference),
    sd_old = sd(Old),
    sd_young = sd(Young),
    sd_diff = sd(Difference)
  )
```

g. Because the data are obtained from the same group of objectives, the variables are now dependent, breaking the assumption of independence of the previous test.

**h.** A hypothesis test using critical value can be used.

**i.** 

**Hypothesis:** $H_0: \mu=0$ vs $H_1: \mu \neq 0$

**Assumptions:** $X_i$ are iid varlables and follow $N(\mu,\sigma^2)$.

**Test statistic:** $T=\frac{\overline{X}-\mu_{0}}{S/\sqrt{n}}$. Under $H_0$, T~$t_{n-1}$.

**Observed test statistic:** $t_0=\frac{1.93-0}{4.18/\sqrt{30}}=0.084$
```{r}
(statistics$mean_diff - 0)/(statistics$sd_diff * sqrt(statistics$n))

qt(c(0.95),30)
```

**Decision:** the observed test statistic $t_0=0.084$ is smaller than 1.6972, so we reject $H_0$ and conclude that there is difference of satisfaction between old adults and younger adults.

**j.**

```{r}
t.test(paired_dat$Difference, mu = 0, conf.level = 0.95)
```

