---
title: "lab 9"
author: "Jing Yang \ 480110301"
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
message = FALSE
warning = FALSE
```

# Questions

## Pain thresholds

```{r}
library(tidyverse)
pain = read_tsv("https://raw.githubusercontent.com/DATA2002/data/master/blonds.txt")
glimpse(pain)
```

```{r}
pain = pain %>% mutate(
  HairColour = factor(HairColour, levels = c("LightBlond", "DarkBlond", "LightBrunette", "DarkBrunette"))
)
levels(pain$HairColour)
```

```{r}
ggplot(pain, aes(x = HairColour, y = Pain)) + 
  geom_boxplot() + theme_classic()
```

```{r}
pain_sum = pain %>% group_by(HairColour) %>% 
  summarise(
    n = n(),
    ybar = mean(Pain),
    
  )
pain_sum
```

```{r}
ni = pain_sum %>% pull(n)
ybar_i = pain_sum %>% pull(ybar)
```

```{r}
pain_aov = aov(Pain ~ HairColour, data = pain)
anova(pain_aov)
```

1. Compute the standard error of each pairwise difference

$\hat \sigma \sqrt{\sum_i \frac{c_i^2}{n_i}}$

```{r}
resid_ms = anova(pain_aov)[2,3]
resid_se = sqrt(resid_ms)
(se.55 =resid_se*sqrt(1/5+1/5))
(se.45 =resid_se*sqrt(1/4+1/5))
```

```{r}
diff_mat = outer(ybar_i, ybar_i, "-")
colnames(diff_mat) = c("LightBlond","DarkBlond","LightBrunette","DarkBrunette")
rownames(diff_mat) = c("LightBlond","DarkBlond","LightBrunette","DarkBrunette")
diff_mat
```

```{r}
se_mat = resid_se * sqrt(outer(1/ni,1/ni,"+"))


t_mat = diff_mat/se_mat
colnames(t_mat) = c("LightBlond","DarkBlond","LightBrunette","DarkBrunette")
rownames(t_mat) = c("LightBlond","DarkBlond","LightBrunette","DarkBrunette")
t_mat
```

```{r}
df = anova(pain_aov)[2,1]
# df = sum(ni)-4

m = qt(1-0.025/6, df=sum(ni)-4)
m
```

```{r}
m = qt(1-0.005/6, df=sum(ni)-4)
m
```

```{r}
library(emmeans)
pain_em = emmeans(pain_aov, ~HairColour)
bonf = contrast(pain_em, method = "pairwise", adjust = "bonferroni")
plot(bonf)+theme_bw()+geom_vline(xintercept=0)
```

```{r}
tukey = contrast(pain_em,  method = "pairwise",adjust = "tukey")
plot(tukey)+theme_bw()+geom_vline(xintercept=0)

scheffe = contrast(pain_em,method = "pairwise",adjust = "scheffe")
plot(scheffe)+theme_bw()+geom_vline(xintercept=0)
```

## Tablet

```{r}
library(tidyverse)
tablet = read_tsv("https://raw.githubusercontent.com/DATA2002/data/master/tablet1.txt")

tabdat = gather(tablet, key = "lab", value = "measurement")
glimpse(tabdat)
```

```{r}
tabdat %>% 
  ggplot() + 
  aes(x = lab, y = measurement, fill = lab) + 
  geom_boxplot() + 
  theme_classic() + 
  labs(y = "Chlorpheniramine maleate (mg)",
       x = "Lab", fill = "")
```

1. Is the mean level of chlorpheniramine maleate in tablets from Lab 1 different from 4.0? State the null hypothesis.

$H_0: \mu_1=4$ 
```{r}
t.test(tablet$Lab1,mu=4)
```


2. Is the mean level of chlorpheniramine maleate in tablets from Lab 1 different from that from Lab 3?

```{r}
t.test(tablet$Lab1,tablet$Lab3)
```

3. Perform a one-way ANOVA to test if the mean levels of chlorpheniramine maleate differ across the seven labs.

```{r}
tablet_aov = aov(measurement ~ lab, data = tabdat)
summary(tablet_aov)
```

4. Obtain a QQ plot of the residuals and comment on the validity of the ANOVA assumptions.

```{r}
library(ggfortify)
autoplot(tablet_aov, which = c(1:2)) + theme_classic()
```

The data is not normally distributed as the dots in the QQ plot does not strictly follow the straight lilne.

5. Perform pairwise post hoc tests to determine which labs are significantly different.

```{r}
tablet_em = emmeans(tablet_aov, ~lab)
conf = confint(pairs(tablet_em))
plot(conf)+theme_bw()+geom_vline(xintercept=0)
```

6. Use a rank based approach to testing for a difference between the means of the 7 labs.
```{r}
tablet_rank = tabdat %>% ungroup() %>% mutate(rank = rank(measurement))
ggplot(tablet_rank, aes(x = factor(lab), y = rank)) +
geom_boxplot() + coord_flip() + theme_classic(base_size =24)
```

```{r}
tablet_kw = aov(rank ~ factor(lab), data = tablet_rank)
broom::tidy(tablet_kw)
```

```{r}
(test = kruskal.test(measurement ~ factor(lab), data = tabdat))
t0 = test$statistic
pval = test$p.value
```

```{r}
# permutation for non-normal distributed data

set.seed(1)
B=2000
lab = tabdat$lab
measurement = tabdat$measurement
kw_stat = vector("numeric",length=B)
for(i in 1:B){
  aov_rank = aov(rank(sample(measurement)) ~ factor(lab))
  kw_stat[i] = broom::tidy(aov_rank)$statistic[1]
}

mean(kw_stat >=t0)
```

